---
title: "Executing NHANES analysis using the PIC-SURE API"
author: "Chirag J Patel and Michael T McDuffie"
date: "May 19, 2016"
output: html_document
---

In this guide, we will show how to estiamte the distribution of lead in children using the NHANES data via the PIC-SURE API.

##Required Libraries for communication
```{r}
library(httr)
library(jsonlite)
library(XML)
```

##Build Base Query Strings for interacting with IRCT

```{r}
IRCT_REST_BASE_URL <- "http://bd2k-picsure.hms.harvard.edu/"

#REST URL
IRCT_CL_SERVICE_URL <- paste(IRCT_REST_BASE_URL,"IRCT-CL/rest/",sep="")

#Service URLS
IRCT_RESOURCE_BASE_URL <- paste(IRCT_CL_SERVICE_URL,"resourceService/",sep="")
IRCT_QUERY_BASE_URL <- paste(IRCT_CL_SERVICE_URL,"queryRESTService/",sep="")
IRCT_RESULTS_BASE_URL <- paste(IRCT_CL_SERVICE_URL,"resultService/",sep="")

#List resources
IRCT_LIST_RESOURCE_URL <- paste(IRCT_RESOURCE_BASE_URL,"resources",sep="")

#Query
IRCT_START_QUERY_URL <- paste(IRCT_QUERY_BASE_URL,"startQuery",sep="")
IRCT_WHERE_QUERY_URL <- paste(IRCT_QUERY_BASE_URL,"whereClause",sep="")
IRCT_RUN_QUERY_URL <- paste(IRCT_QUERY_BASE_URL,"runQuery",sep="")

#Add Select Clause
IRCT_SELECT_QUERY_URL <- paste(IRCT_QUERY_BASE_URL,"selectClause",sep="")

#Get Results
IRCT_GET_JSON_RESULTS_URL <- paste(IRCT_RESULTS_BASE_URL,"download/json",sep="")
```

##Start a conversation
```{r}
response <- GET(IRCT_START_QUERY_URL, verbose())
conversationId <- content(response)$cid
```

##Add a where clause to filter patient list
```{r}
whereParameterList <- list(type="where", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/SEX/", 
                            logicalOperator="AND",
                            predicate="CONTAINS",
                            "data-encounter"="No",
                            cid=conversationId
                           )
response <- GET(IRCT_WHERE_QUERY_URL, query=whereParameterList, verbose())
content(response)
```

##Add select clauses
```{r}
selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/AGE/",
                            alias="AGE",
                            cid=conversationId)
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)

selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/RACE/",
                            alias="Race",
                            cid=conversationId)
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)

selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/SDDSRVYR/", 
                            alias="Survey Year",
                            cid=conversationId
                           )
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)


selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/SDMVPSU/", 
                            alias="Survey PSU",
                            cid=conversationId
                           )
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)

selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/SDMVSTRA/", 
                            alias="Survey Strata",
                            cid=conversationId
                           )
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)

selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/demographics/WTMEC2YR/", 
                            alias="Survey Weighting",
                            cid=conversationId
                           )
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)



selectParameterList <- list(type="select", 
                            field="NHANES Public/Public Studies///NHANES/NHANES/laboratory/heavy metals/Lead (ug/dL)", 
                            alias="Lead",
                            cid=conversationId
                           )
response <- GET(IRCT_SELECT_QUERY_URL, query=selectParameterList, verbose())
content(response)


```

##Run the full query and store the result id
```{r}
runQueryList <- list(cid=conversationId)
response <- GET(IRCT_RUN_QUERY_URL, query=runQueryList, verbose())
resultId <- content(response)$resultId
content(response)
```

##Retrieve Results
```{r}
response <- GET(paste(IRCT_GET_JSON_RESULTS_URL,'/',resultId,sep=""), verbose())
results <- content(response)
results<- data.frame(do.call("rbind", results))
```

## Compute the quantiles of serum lead distribution of people under the age of 18 by sex in 2005-2006
```{r}
library(survey)
results[, 'X.NHANES.laboratory.heavy.metals.Lead..ug.dL..'] <- as.numeric(results[, 'X.NHANES.laboratory.heavy.metals.Lead..ug.dL..'])
results[, 'X.NHANES.demographics.AGE.'] <- as.numeric(results[, 'X.NHANES.demographics.AGE.'])
results[, 'X.NHANES.demographics.SDMVSTRA.'] <- as.numeric(results[, 'X.NHANES.demographics.SDMVSTRA.'])
results[, 'X.NHANES.demographics.SDDSRVYR.'] <- as.numeric(results[, 'X.NHANES.demographics.SDDSRVYR.'])
results[, 'X.NHANES.demographics.WTMEC2YR.'] <- as.numeric(results[, 'X.NHANES.demographics.WTMEC2YR.'])
results[, 'X.NHANES.demographics.SDMVPSU'] <- as.numeric(results[, 'X.NHANES.demographics.SDMVPSU'])
dsn <- svydesign(id=~X.NHANES.demographics.SDMVPSU, strata=~X.NHANES.demographics.SDMVSTRA., probs=~X.NHANES.demographics.WTMEC2YR., nest=T,data=subset(dat, X.NHANES.demographics.WTMEC2YR. > 0)) # first cret a survey design object, speciifying the sampling units (SDMVPSU), the strata (SDMVSTRA), and probability weight of being selected WTMEC2YR
dsn.2005 <- subset(dsn, X.NHANES.demographics.SDDSRVYR. == 4)
svyquantile(~X.NHANES.laboratory.heavy.metals.Lead..ug.dL.., quantiles= c(0, .25,.5,.75, 1), dsn.2005, na.rm=T)
```