---
title: "NHANES Scientific Data User Guide"
author: "Chirag J Patel"
date: "May 19, 2016"
output: html_document
---


This presents how to use to how to use the NHANES datasets and to make inferences that account for the non-random and stratified nature of the survey.
Comments to:
Chirag Patel (chirag@hms.harvard.edu)
5/19/2016


Load the .Rdata object
```{r}
load('./nh_99-06.Rdata')
```

How do we figure out what is a exposure and what is a phenotype in NHANES?
Hint: Use the 'VarDescription' data.frame:
```{r}
head(VarDescription) ## this gives the variable name and description and broad category for each variable (called 'var_desc_ewas')

as.data.frame(table(VarDescription$category)) ##  the types of variables

```

Next, how does survey-weighted regression work?
Suppose we want to look at the association between fasting glucose and BMI (adjusted by age and sex) in the 2003-2004 survey.

Under a normal study sample, we would simply use lm:
```{r}
dat <- subset(MainTable, SDDSRVYR == 3) # subset for 2003-2004
mod <- lm(LBXGLU ~ BMXBMI + RIDAGEYR + female, dat)
summary(mod)
```

But with NHANES, this is technically not correct. We need to use survey-weighting to accomodate the survey sampling of the data:
```{r}
library(survey)
dsn <- svydesign(id=~SDMVPSU, strata=~SDMVSTRA, probs=~WTMEC2YR, nest=T,data=subset(dat, WTMEC2YR > 0)) # first cret a survey design object, specififying the sampling units (SDMVPSU), the strata (SDMVSTRA), and probability weight of being selected WTMEC2YR
mod.svy <- svyglm(LBXGLU  ~ BMXBMI + RIDAGEYR + female, design=dsn) ## now use SVYGLM; 
summary(mod.svy)  #slightly different estimates
```

Lets try logistic regression, looking at the clinical diagnosis of diabetes (LBXGLU >= 126) using logistic regression:
```{r}
mod.svy.t2d <- svyglm(I(LBXGLU >=125)  ~ BMXBMI + RIDAGEYR + female, design=dsn, family=quasibinomial()) #depending on the family= parameter, you can use this for logistic regression, as well.
summary(mod.svy.t2d) #t2d increases by 10% per 1 unit increase in BMI.
```

What about survival analysis? Different beast! In survival analyses, we require whther the person died at the time of querying survival (0 or 1), and time to querying (e.g., 1 month, 5 months, etc)
These are coded as MORTSTAT and PERMTH_EXM (time to death from the the examination survey) in the MainTable respectively.
For a 1 unit glucose increase, what is the hazard of death adjusting for age and sex for participants surveyed in 1999-2000?
```{r}
suvdat <- subset(MainTable, !is.na(MORTSTAT) & !is.na(PERMTH_EXM) & SDDSRVYR == 1) ## get all data from 1999-2000
dsn <- svydesign(id=~SDMVPSU, strata=~SDMVSTRA, probs=~WTMEC2YR, nest=T,data=suvdat)
mod.cox.svy <- svycoxph(Surv(PERMTH_EXM, MORTSTAT) ~ RIDAGEYR + female + LBXGLU, dsn)
summary(mod.cox.svy)
```

With those basics in hand, now we can tackle executing exposome-like associations in NHANES.

For example, we recently found an association between serum cadmium and mortality (see: Patel CJ et al, Int. J. of Epidem., 2013). Serum cadmium was found to be signficantly associated with all-cause mortality.

What is the variable name for "cadmium?"
```{r}
VarDescription[grep('cadmium', VarDescription$var_desc, ignore.case = T), ] # looks like it is LBXBCD
```

```{r}
suvdat <- subset(MainTable, !is.na(MORTSTAT) & !is.na(PERMTH_EXM) & SDDSRVYR == 1) ## get all data from 1999-2000
dsn <- svydesign(id=~SDMVPSU, strata=~SDMVSTRA, probs=~WTMEC2YR, nest=T,data=suvdat)
mod.cox.cadmium.1 <- svycoxph(Surv(PERMTH_EXM, MORTSTAT) ~ RIDAGEYR + female + LBXBCD, dsn)
summary(mod.cox.cadmium.1)
```

Does it replicate in the 2001-2002 survey?
```{r}
suvdat <- subset(MainTable, !is.na(MORTSTAT) & !is.na(PERMTH_EXM) & SDDSRVYR == 2) ## 2003-2004 survey
dsn <- svydesign(id=~SDMVPSU, strata=~SDMVSTRA, probs=~WTMEC2YR, nest=T,data=suvdat)
mod.cox.cadmium.2 <- svycoxph(Surv(PERMTH_EXM, MORTSTAT) ~ RIDAGEYR + female + LBXBCD, dsn) # yes, strong association in 2001-2002 as well
summary(mod.cox.cadmium.2)
```

